---
title: "Playing with Spark"
output: html_notebook
---

```{r}
###########################
#####  DATABASE.Rmd  ######
###########################
```

```{r include=FALSE, echo=FALSE}
options(rsparkling.sparklingwater.version = "2.0.3")
library(rsparkling)
library(sparklyr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)
library(plotly)
library(dygraphs)

source("helper.R")
```

```{r}
sc <- spark_connect(master = "local", version = "2.0.0", hadoop_version="2.7")
crimes_tbl <- spark_read_csv(sc, "crimes", "./data/rows.csv")

src_tbls(sc)
dim(crimes_tbl)
```
Make spark table with only the columns we need
```{r include = FALSE}
crimes_tbl <- crimes_tbl %>%
  mutate(Date2 = substr(Date, 1, 10), 
         Month = as.integer(substr(Date, 1,2)), 
         Day = as.integer(substr(Date, 4,2)), 
         Year2 = as.integer(substr(Date, 7,4)), 
         Time_of_Day = substr(Date, 12, 8), 
         AM_PM = substr(Date, 21, 2)) %>%
  mutate(Hour = #transform time_of_Day to be numeric hours.fractional_hours
           as.numeric(substr(Time_of_Day, 1, 2)) + #hours
           (as.numeric(substr(Time_of_Day, 4, 2)) / 60) + #minutes to fractionsl hours
           (as.numeric(substr(Time_of_Day, 7, 2)) / 60 / 60) #seconds to fractional hours
         ) %>% 
  mutate(Time_of_Day = if_else((AM_PM == "PM" && floor(Time_of_Day) < 12) || 
                                 (AM_PM == "AM" && floor(Time_of_Day) == 12), 
                               (Time_of_Day + 12) %% 24.0, Time_of_Day)) %>%
  select(Primary_Type, Latitude, Longitude, Date, Month, Day, Year, Year2, Time_of_Day, Hour) 

crimes_tbl
```
# SETUP
#===========
# POPULATION DENSITY CALCULATION
# ===============================
```{r}
# get min and max year 
yearRange <- crimes_tbl %>% select(Year) %>% collapse() %>% summarise(min = min(Year), max = max(Year)) %>% select(min, max) %>% collect()

population_2000 <- 2893666 #Values obtained from census data
population_2010 <- 2695598

# CAGR - Compound Annual Growth Rate of the Population
cagr <- (population_2010/population_2000)^(1/(10*12)) - 1

num_years <- yearRange$max - yearRange$min
num_months <- num_years*12

year <- rep(yearRange$min:yearRange$max, each=12)
month <- rep(1:12, num_years)

pop <- population_2000

for (i in 1:(num_months-1)) {
	pop[i+1] <- pop[i] * (1+cagr)	
}	


df.population <- sdf_copy_to(sc, x = data.frame(Year=year, Month=month, Population=pop/100000), name = "population", overwrite = TRUE)
df.population
```
for compatibility between code
```{r}
df.raw <- crimes_tbl
```
# AREA OF CHICAGO
# ===============================
```{r}
a <- crimes_tbl %>% summarise(minLat = min(Latitude), maxLat = max(Latitude), minLong = min(Longitude), maxLong = max(Longitude)) %>% collect()
a
area_chicago <- (a$maxLat - a$minLat) * (a$maxLong - a$minLong)
```

# DATA PREPARATION
# ======================
```{r}

# Inputs:
#	loc - a list containing 2 elements: vector(latitude/longitude) for bottom-left and top-right points
#	tim - a list containing 2 elements: vector(start and end time), interval (Y, M)
#	typ - a list containing 1 element:  vector(primary type)
tim <- list(year=c(yearRange$min, yearRange$max), month=c(1,2))
typ <- list(prm=c("HOMICIDE", "ASSAULT", "BATTERY", "NARCOTICS"))
loc <- list(bl = c(a$minLat, a$minLong),
            tr = c(a$maxLat, a$maxLong))

# typ <- reactive({
#   ret <- list(prm = input$tab2_)
# })
# loc <- reactive({
#   bounds <- input$tab2_map_bounds
#   if(is.null(bounds)) {
#     ret <- list(bl = c(min(df.raw$Latitude, na.rm=T), min(df.raw$Longitude, na.rm=T)),
#                 tr = c(max(df.raw$Latitude, na.rm=T), max(df.raw$Longitude, na.rm=T)))
#   } else {
#     ret <- list(bl = c(bounds$south, bounds$west), tr = c(bounds$north, bounds$east))
#   }
#   return(ret)
# })
```
# Determine the ratio of the area selected
#-----------------------
```{r}
sel_area <- (loc$tr[1]-loc$bl[1])*(loc$tr[2]-loc$bl[2]); sel_area
sel_ratio <- sel_area/area_chicago
```
# Subset Data based on input parameters
#-----------------------
```{r}

df <- df.raw %>%
	filter(!is.na(Latitude), !is.na(Longitude), !is.na(Date), !is.na(`Primary_Type`)) %>%
	filter(`Primary_Type` %in% typ$prm) %>%
  filter(Latitude >= loc$bl[1], Latitude <= loc$tr[1], Longitude >= loc$bl[2], Longitude <= loc$tr[2]) %>%
	filter(((Year>tim$year[1] & Year<tim$year[2]) | (Year==tim$year[1] & Month>=tim$month[1]) | (Year==tim$year[2] & Month<=tim$month[2]))) %>%
# Join with calcualted population data
# -----------------------
	inner_join(df.population, by = c("Year", "Month"))
df
```
# Calculate the Crime Rate based on Area and Population Density 
# -----------------------
```{r}

df.totals <- dfr%>%
	group_by(Year, Month, Population) %>%
  summarise(Total = n()) %>%
  collapse() %>%
  mutate(CrimeRate = Total/(Population*sel_ratio)) %>% #rate / 100,000
  compute()

df.totals
```
# DESCRIPTIVE STATISTICS
# ===============================
```{r}
# 1. Mean and Standard Dev for CrimeRate (Monthly)
# -------------------------------
a <- df.totals %>% select(CrimeRate) %>% collect()
monthly_mean <- mean(a$CrimeRate); monthly_mean
monthly_std <- sd(a$CrimeRate); monthly_std
#    Mean and Standard Dev for Crime Rate (Annually)
b <- df.totals %>%
	group_by(Year, Population) %>%
  summarise(Total = n()) %>%
  mutate(CrimeRate = Total/(Population*sel_ratio)) %>% #rate / 100,000
  select(CrimeRate) %>%
  collect()

annual_total <- b %>%	summarise(Total=sum(CrimeRate)); annual_total
annual_mean <- mean(annual_total$Total); annual_mean
annual_std <- sd(annual_total$Total); annual_std

#2. Average Annual Growth Rate 
monthly_growthrate <- diff(a$CrimeRate) / a$CrimeRate[1:(length(a$CrimeRate)-1)]; monthly_growthrate
avg_monthly_growthrate <- mean(monthly_growthrate); avg_monthly_growthrate


# 3. Mean Time of Day for the Occurence of Crime
# -------------------------------
hour_stats <- dfr %>%
	summarise(mean_hour=mean(Hour), std_hour=sd(Hour)); hour_stats
```

# DESCRIPTIVE GRAPHS
# ===============================

# 1. Histogram Displaying Frequency of Crime based on Time of Day
# -------------------------------
```{r}
ddd <- as.data.frame(dfr %>% summarise(Hour = floor(Hour)) %>% collect())$Hour; ddd
p1 <- plot_ly(alpha = 0.6) %>% 
  add_histogram(x = df$Hour) %>% 
  layout(xaxis = list(title="Time (Hours)"), yaxis=list(title="Frequency")); p1
```

```{r}
# 2. Time Series Graph Displaying the Number of Incidents of Crime in Every Month
# ------------------------------

# Convert to Time Series
# -------------------------------

df.month <- as.data.frame(df.totals); df.month
ts.month <- df.month %>%
	select(CrimeRate) %>%
	ts( frequency=12, start=c(tim$year[1],tim$month[1]), end=c(tim$year[2],tim$month[2])); ts.month

dygraph(ts.month, main = "Monthly Incidents of Crime in Chicago") %>%
	dyOptions(fillGraph=TRUE, fillAlpha=0.2) %>%
	dySeries(label=paste(typ$prm, sep="", collapse="+")) %>%
	dyAxis("x", label = "Time (Months)", valueRange=c(tim$year[1], tim$year[2])) %>%
	dyAxis("y", label = "Number of Incidents") %>%
	dyHighlight(highlightSeriesOpts = list(strokeWidth = 1)) %>%
	dyRangeSelector()


# INFERENTIAL STATISTICS - FORECASTING
# ===============================
# Decompose Time Series into Trend, Seasonal, and Random components
# -------------------------------
ts.decomp <- decompose(ts.month)
plot(ts.decomp)



```














