---
title: "Playing with Spark"
output: html_notebook
---

```{r}
###########################
#####  DATABASE.Rmd  ######
###########################
```

```{r}
options(rsparkling.sparklingwater.version = "2.0.3")
library(rsparkling)
library(sparklyr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)

source("helper.R")
```

```{r}
sc <- spark_connect(master = "local", version = "2.0.0", hadoop_version="2.7")
crimes_tbl <- spark_read_csv(sc, "crimes", "./data/rows.csv")
src_tbls(sc)
```

Some variables
collect() means that the table is brought back to a locally and stored in loc
```{r}
crimeTypes <- c("ASSAULT", "BATTERY")
timeFrame <- "Y"
loc <- crimes_tbl %>% 
  summarise(blat = min(Latitude), blon = min(Longitude), tlat = max(Latitude), tlon = max(Longitude)) %>% 
  collect()
```

Make a table of just violent crimes
Keep it on the server with compute() violentCrimes points to this table (tbl_spark)
```{r}
violentCrimes <- crimes_tbl %>%
  filter(Primary_Type %in% crimeTypes) %>%
  compute()
head(violentCrimes)
```

```{r}
violentCrimes2 <- violentCrimes %>%
  mutate(Date_time = as.numeric(as.POSIXct(Date, format = "%m/%d/%Y %I:%M:%S %p", tz = "CST6CDT")), 
         Date_date = Date) %>%#split(Date, " ")[1]) %>%# %>% split("/")) %>% #, 
  #collect() %>%
  #summarise(Date_3 = as.Date(Date)) %>%
  select(Date_time, Date_date) %>%
  compute()
head(violentCrimes2)
```

```{r}
%>%
  collapse() %>%
  mutate

           strptime())
  mutate(Year = as.POSIXct(as.character(Date),format="%m/%d/%Y %I:%M:%S %p", tz="CST6CDT")
         )

# create a table on the server
# 
violentCrimes <- crimes_tbl %>%
  mutate(assault = ifelse(Primary_Type == "ASSAULT", ))
  

  

applicants_tbl %>%
  mutate(male = ifelse(sex == "M", n_all, 0), female = ifelse(sex == "F", n_all, 0)) %>%
  group_by(year) %>%
  summarize(Male = sum(male) / 1000000, Female = sum(female) / 1000000) %>%
  arrange(year) %>%
  collect

birthsYearly %>%
  dygraph(main = "Total US Births (SSN)", ylab = "Millions") %>%
  dySeries("Female") %>%
  dySeries("Male") %>%
  dyOptions(stackedGraph = TRUE) %>%
  dyRangeSelector(height = 20)


head(crimes_tbl$Latitude)

tim <- list(pd = c(2003, 2016), int = "Y")
typ <- list(prm = c("ASSAULT", "BATTERY"))
loc <- crimes_tbl %>% summarise(blat = min(Latitude), blon = min(Longitude), tlat = max(Latitude), tlon = max(Longitude)) %>% collect()
loc



df <- crimes_tbl %>%
  select(Latitude, Longitude, Date, Primary_Type) %>%
  filter(Primary_Type %in% typ$prm) %>%
  arrange(Primary_Type) %>%
  compute()

df

crimes_tbl %>% filter(Primary_Type %in% c("ASSAULT", "BATTERY")) %>% group_by(Primary_Type)
  




# Subset by type
# ------------------
df <- df[df$Primary.Type %in% typ$prm,]
# Subset by location
# ------------------
#df <- df[df$Latitude >= loc$bl[1] 
 #        & df$Latitude <= loc$tr[1]
  #       & df$Longitude >= loc$bl[2]	
   #      & df$Longitude <= loc$tr[2],]
# Subset by time
# ------------------
df$Date <- as.POSIXct(as.character(df$Date),format="%m/%d/%Y %I:%M:%S %p", tz="CST6CDT")

df$Year <- year(df$Date)
df$Month <- month(df$Date)
df$Day <- day(df$Date)
df$Hour <- hour(df$Date)
df <- switch(tim$int,
             "Y" = df[df$Year>=tim$pd[1] & df$Year<=tim$pd[2],],
             "M" = df[df$Month>=tim$pd[1] & df$Month<=tim$pd[2],],
             "D" = df[df$Day>=tim$pd[1] & df$Day<=tim$pd[2],])




df.grp <- df %>% group_by(Hour)
  
birthsYearly <- crimes_tbl %>%
  mutate(assault = )
  
  
  applicants_tbl %>%
  mutate(male = ifelse(sex == "M", n_all, 0), female = ifelse(sex == "F", n_all, 0)) %>%
  group_by(year) %>%
  summarize(Male = sum(male) / 1000000, Female = sum(female) / 1000000) %>%
  arrange(year) %>%
  collect
 
birthsYearly %>%
  dygraph(main = "Total US Births (SSN)", ylab = "Millions") %>%
  dySeries("Female") %>%
  dySeries("Male") %>%
  dyOptions(stackedGraph = TRUE) %>%
  dyRangeSelector(height = 20) 





```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
